---
title: "Data_Analysis_LE"
author: "Elizabeth Newman and Lili Maycock"
date: "2026-09-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We decided that we wanted to model the probability of receiving professional sanctions based on the other variables available in the dataset.  


```{r}
library(readr)
library(ggplot2)
hallucination_cases <- read_csv("hallucination_cases.csv")
head(hallucination_cases)
```


```{r}
#checking types of parties 
library(dplyr)
hallucination_cases %>%
  count(`Party(ies)`)
```
```{r}
#checking legal fields 
hallucination_cases %>%
  count(`Legal Field Primary`)
```
```{r}
#checking what the distribution of professional sanction vs no professional sanction
hallucination_cases %>%
  count(`Professional Sanction`)
```
```{r}
#checking what the distribution of professional sanction vs no professional sanction
hallucination_cases %>%
  count(`State(s)`)
```
Because we can see that there a lot of countries with only one case, it doesn't make to make every country a variable. That would add 45 predictors to the model. We will thus make a variable for all states with 50 or more cases and other: so Australia, Canada, Israel, UK, the US and all other countries. 

```{r}
hallucination_cases <- hallucination_cases |>
  filter(!is.na(`State(s)`)) |>
  mutate(
    country_group = if_else(
      `State(s)` %in% c("Australia", "Canada", "Israel", "UK", "United States"),
      `State(s)`,
      "Other"
    )
  )
```

Looking at the where we get the data from, we can see that the variable AI tool contains "implied" which is when AI usage was evident but not admitted. 

```{r}
#checking what the distribution of professional sanction vs no professional sanction
hallucination_cases %>%
  count(`AI Tool`)
```
Similarly to country, but more extreme, we have 104 options for AI tool, which is too many variables. We can thus separate this into three categories: Implied, Unidentified, and Specific AI Tool. 

```{r}
hallucination_cases <- hallucination_cases |>
  filter(!is.na(`AI Tool`)) %>%
  mutate(
    ai_tool_group = case_when(
      `AI Tool` == "Unidentified" ~ "Unidentified",
      `AI Tool` == "Implied" ~ "Implied",
      `AI Tool` == "implied" ~ "Implied",
      TRUE ~ "Specific AI Tool"
    )
  )
```


```{r}
ggplot(hallucination_cases, aes(x = `Legal Field Primary`, fill = `Professional Sanction`)) +
  geom_bar(position = "fill") +
  labs(
    x = "Primary Legal Field",
    y = "Proportion of Cases",
    fill = "Professional Sanctions"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(hallucination_cases, aes(x = `country_group`, fill = `Professional Sanction`)) +
  geom_bar(position = "fill") +
  labs(
    x = "Country",
    y = "Proportion of Cases",
    fill = "Professional Sanctions"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(hallucination_cases, aes(x = `ai_tool_group`, fill = `Professional Sanction`)) +
  geom_bar(position = "fill") +
  labs(
    x = "AI Tool Admitted",
    y = "Proportion of Cases",
    fill = "Professional Sanctions"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
library(broom)
#making professional sanctions a binart variable and then running a glm 
hallucination_cases <- hallucination_cases %>%
  mutate(
    professional_sanction_binary = if_else(
      `Professional Sanction` == "Yes",
      1,
      0
    )
  )
logistic_model <- glm(`professional_sanction_binary` ~ `country_group` + `Legal Field Primary` + `ai_tool_group`, data = hallucination_cases, family = "binomial") 
summary(logistic_model)
#tidy(logistic_model)
```


